---
title: "EDA Continued"
author: "DSCI 522 Group 304"
date: "2020-01-22"
output: gitub_document
---

This EDA file looks at only whether there is still a difference in scores between Aboriginal and Non Aboriginal students using a subset of schools that must have both Aboriginal and Non Aboriginal students. 

The results shown below indicate that there is still a significant difference in scores between the subgroups even when looking only at schools that have both Aboriginal and Non Aboriginal students. 

```{r}
library(docopt)
library(readr)
library(tidyverse)
```


These first 3 cells show that there's a difference in mean scores between subgroups even if the dataset was filtered to only include schools with both Aboriginal and Non Aboriginal students.

```{r Difference in Scores for Numeracy Test}
data <- read_csv('../../data/filtered_schools_both_subgroups.csv')

# Filters data for only the subgroups of interest
filtered_data = data %>%
  filter(sub_population == 'ABORIGINAL' | sub_population == 'NON ABORIGINAL') %>%
  filter(fsa_skill_code == 'Numeracy')

t.test(score ~ sub_population, data = filtered_data, var.equal = FALSE)
```
```{r Difference in Scores for Reading Test}

# Filters data for only the subgroups of interest
filtered_data = data %>%
  filter(sub_population == 'ABORIGINAL' | sub_population == 'NON ABORIGINAL') %>%
  filter(fsa_skill_code == 'Reading')

t.test(score ~ sub_population, data = filtered_data, var.equal = FALSE)
```

```{r Difference in Scores for Writing Test}

# Filters data for only the subgroups of interest
filtered_data = data %>%
  filter(sub_population == 'ABORIGINAL' | sub_population == 'NON ABORIGINAL') %>%
  filter(fsa_skill_code == 'Writing')

t.test(score ~ sub_population, data = filtered_data, var.equal = FALSE)
```

The following code cells check using Tiffany's way: comparing scores between schools with both subgroups and schools with only one subgroup (Non Aboriginal). These results show a significant difference between the scores.


```{r Difference in Scores for Numeracy Test}
data <- read_csv('../../data/new_clean_data.csv')

# Filters data for only the subgroups of interest
filtered_data = data %>%
  filter(sub_population == 'ALL STUDENTS') %>%
  filter(fsa_skill_code == 'Numeracy')

t.test(score ~ has_both, data = filtered_data, var.equal = FALSE)
```

```{r Difference in Scores for Reading Test}

# Filters data for only the subgroups of interest
filtered_data = data %>%
  filter(sub_population == 'ALL STUDENTS') %>%
  filter(fsa_skill_code == 'Reading')

t.test(score ~ has_both, data = filtered_data, var.equal = FALSE)
```

```{r Difference in Scores for Writing Test}

# Filters data for only the subgroups of interest
filtered_data = data %>%
  filter(sub_population == 'ALL STUDENTS') %>%
  filter(fsa_skill_code == 'Writing')

t.test(score ~ has_both, data = filtered_data, var.equal = FALSE)
```


